---
title: "How did the opening of the Silver Line of the Washington Metro System impact adjacent rental prices?"
format: gfm
editor: visual
---

Step 1. Install necessary packages.

```{r}
install.packages("tidyverse")
install.packages("kableExtra")
```

Step 2. Declare that you will use these packages in this session.

```{r}
library("tidyverse")
library("kableExtra")
```

Step 3. Upload the dataframe that you have created in Spring 2024 into the repository.

Step 4. Open the dataframe into the RStudio Environment.

```{r}
df<-read.csv("panel.csv")

df2<-df %>%
  mutate(date = as.Date(Date, format = "%Y.%m.%d"), month = month(date), year = year(date))
```

Step 5. Use the **head** and **kable** function showcase the first 10 rows of the dataframe to the reader.

```{r}
kable(head(df2))
```

## Question 1: What is the frequency of this data frame?

Answer: Monthly

## Question 2: What is the cross-sectional (geographical) unit of this data frame?

Answer: Zip code

Step 6. Use the **names** function to display all the variables (column) in the dataframe.

```{r}
names(df2)

```

## Question 3: Which column represents the treatment variable of interest?

Answer: silver

## Question 4: Which column represents the outcome variable of interest?

Answer: ZORI

Step 7: Create a boxplot to visualize the distribution of the outcome variable under treatment and no treatment.

```{r}
#| eval: false
ggplot(df2, aes(x=ZORI)) +
  geom_histogram() +
  facet_wrap(~silver)
```

Step 8: Fit a regression model $y=\beta_0 + \beta_1 x + \epsilon$ where $y$ is the outcome variable and $x$ is the treatment variable. Use the **summary** function to display the results.

```{r}
model1<-lm(ZORI ~ silver, data=df2)

summary(model1)
```

## Question 5: What is the predicted value of the outcome variable when treatment=0?

Answer: 2040.129

## Question 6: What is predicted value of the outcome variable when treatment=1?

Answer: 2040.129 + 188.653 = 2228.782

## Question 7: What is the equation that describes the linear regression above? Please include an explanation of the variables and subscripts.

Answer: $$ZORI = \beta_0 + \beta_1silver + \epsilon$$

## Question 8: What fixed effects can be included in the regression? What does each fixed effects control for? Please include a new equation that incorporates the fixed effects.

Answer: $$ZORI = \beta_0 + \beta_1 silver + \beta_2 city + \beta_3 zip\_code + \beta_4 month + \beta_5 year + \epsilon$$

## Question 9: What is the impact of the treatment effect once fixed effects are included?

```{r}
model2 <- lm(ZORI ~ silver + as.factor(City) + as.factor(ZIPCODE) + as.factor(month) + as.factor(year), data=df2)
summary(model2)
```

## 

Answer: When fixed effects are included, the silver line's effect jumps from 188 to 630.

# Questions for Week 5

## Question 10: In a difference-in-differences (DiD) model, what is the treatment GROUP?

Answer:

## Question 11: In a DiD model, what are the control groups?

Answer:

## Question 12: What is the DiD regression equation that will answer your research question?

## Question 13: Run your DiD regressions below. What are the results of the DiD regression?

## Question 14: What are the next steps of your research?

Step 9: Change the document format to gfm

Step 10: Save this document as README.qmd

Step 11: Render the document. README.md file should be created after this process.

Step 12: Push the document back to GitHub and observe your beautiful document in your repository!

Step 13: If your team has a complete dataframe that includes both the treated and outcome variable, you are done with the assignment. If not, make a research plan in Notion to collect data on the outcome and treatment variable and combine it into one dataframe.
